---
title: "Experiment 3: Orientation with lower constrast"
format: html
---

## Libraries and settings required for analysis
```{r setup}
library(tidyverse)

library(cmdstanr)
library(fs)
library(glue)
library(kilter) # devtools:install_github("alexander-pastukhov/kilter", dependencies=TRUE)
library(loo)
library(patchwork)

source("utils.R")
```


## Data import

```{r import data}
results <-
  # get names for all csv files 
  fs::dir_ls("Data/exp03-orientation-low-contrast", glob="*.csv") |>
  
  # load them one at a time specifying the column structure
  purrr::map(~read_csv(., col_types = cols(
    Participant = col_character(),
    Session = col_character(),
    Block = col_integer(),
    Task = col_character(),
    Order = col_character(),
    ComplementaryOri = col_logical(),
    Trial = col_integer(),
    OnsetDelay = col_double(),
    Orientation = col_double(),
    Color = col_character(),
    ColorIsTarget = col_logical(),
    ColorResponse = col_character(),
    ColorRT = col_double(),
    OrientationResponse = col_double(),
    OrientationRT = col_double()
  ))) |>
  
  # bind into a single table
  list_rbind() |>
  
  # block 1: single central task practice
  filter(Block > 1) |>

  # Mark first trial for each participant (no prior history)
  mutate(IsFirstTrial = Block ==2 & Trial == 1) |>

  # turning selected variables into factors
  mutate(Task = factor(Task, levels = c("single", "dual")),
         Order = factor(Order, levels = c("ascending", "random", "descending")),
         Participant = factor(Participant, labels = c(1:12))) |>
  
  # computing whether response on central task was correct
  mutate(ColorTaskCorrect = (ColorIsTarget & ColorResponse == "target") | (!ColorIsTarget & ColorResponse ==  "distractor")) |>
  
  # adjusting for circularity and symmetry of orientation response
  mutate(Orientation = ifelse(ComplementaryOri, -Orientation, Orientation),
         OrientationResponse = ifelse(ComplementaryOri, -OrientationResponse, OrientationResponse),
         DeltaOri = OrientationResponse - Orientation,
         OriResponse = case_when(DeltaOri > 90 & OrientationResponse > 180  ~ OrientationResponse - 180,
                                 DeltaOri < 90 & OrientationResponse < -180 ~  OrientationResponse + 180,
                               TRUE ~ OrientationResponse),
         ResponseError = OriResponse - Orientation) |>
  rename(RawOrientationResponse = OrientationResponse)
```

## Statistical analysis

Preparing data for Stan
```{r}
stan_data <- list(
  DataN = nrow(results),
  TaskN = nlevels(results$Task),
  ParticipantsN = nlevels(results$Participant),

  Ori = results$Orientation,
  Response = results$OriResponse,
  Task = as.integer(results$Task),
  Participant = as.integer(results$Participant),
  IsFirstTrial = results$IsFirstTrial,
  Ori_max = 65,
  Ori_min = 25
)
```

Compiling models
```{r}
model_names <- c("bayesian_integration", "ubi_ur", "ubi_r", "ubi_re", "rect_lu")
models <- map(model_names, ~cmdstan_model(glue("Stan/ori_{.}.stan")))
names(models) <- model_names
```

Sampling
```{r}
for(a_model in model_names) {
  cat(a_model, "\n")
  loo_name <- glue("LOO/loo-exp3-{a_model}.RDS")
  draws_with_predictions_name <- glue("DrawsWithPredictions/draws-exp3-{a_model}.RDS")
  draws_name <- glue("Draws/draws-exp3-{a_model}.RDS")
  
  if (fs::file_exists(loo_name) & fs::file_exists(draws_with_predictions_name)) next;
  
  fit <- models[[a_model]]$sample(stan_data, chains = 4, parallel_chains = future::availableCores())
  fit_loo <- fit$loo()
  saveRDS(fit_loo, loo_name) 
  
  draws <- fit$draws(format = "df") |> as_tibble() |> select(-starts_with("log_lik"))
  saveRDS(draws, draws_with_predictions_name)
  
  draws_without_predictions <- draws |> select(starts_with(c("mu_params", "l_rho_params", "sigma_params", "sigma_params", "z_params", "sigma_task")))
  saveRDS(draws_without_predictions, draws_name)
}
```


Comparing models via leave-one-out information criterion
```{r}
loos <- purrr::map(model_names, ~readRDS(glue("LOO/loo-exp3-{.}.RDS")))
names(loos) <- model_names
summarize_loo_comparison(loos)
saveRDS(summarize_loo_comparison(loos), "LOO/loo-df-exp3.RDS")
```


## Generating summary for predictions at the group level for absolute orientation

```{r}
draws <- readRDS("DrawsWithPredictions/draws-exp3-rect_lu.RDS")
mu <- kilter::extract_stan_posterior_matrix(draws, "mu")
```

```{r}
# Bootstrapping behavioral averages
bootstrapped_abs <- bootstrap_group_averages("BootstrappedAverages/exp3-absolute-rect_lu.RDS", results, Orientation, OriResponse, compute_average_absolute_response, order = FALSE)

# Posterior predictions
posterior_abs <- posterior_group_averages_from_mu("PosteriorAverages/exp3-absolute-rect_lu.RDS", results, mu, Orientation, OriResponse, compute_average_absolute_response, order = FALSE)

plot <- plot_model_predictions("rect_lu", posterior_abs, bootstrapped_abs, Orientation, Response) +
  geom_abline(aes(intercept = 0, slope = 1), linetype = "longdash", linewidth = 0.7)
saveRDS(plot, file = glue("ParameterPlots/abs-posterior-prediction-exp03-rect_lu.RDS"))
plot
ggsave("Figures/posterior-abs-exp03_rect_lu.png", units = "cm", width = 16, height = 10)
ggsave("Figures/posterior-abs-exp03_rect_lu.svg", units = "cm", width = 16, height = 10)
```
### per order
```{r}
# Bootstrapping behavioral averages
bootstrapped_abs <- bootstrap_group_averages("BootstrappedAverages/exp3-absolute-rect_lu-order.RDS", results, Orientation, OriResponse, compute_average_absolute_response, order = TRUE)

# Posterior predictions
posterior_abs <- posterior_group_averages_from_mu("PosteriorAverages/exp3-absolute-rect_lu-order.RDS", results, mu, Orientation, OriResponse, compute_average_absolute_response, order = TRUE)

plot <- plot_model_predictions_order("rect_lu", posterior_abs, bootstrapped_abs, Orientation, Response) +
  geom_abline(aes(intercept = 0, slope = 1), linetype = "longdash", linewidth = 0.7)
saveRDS(plot, file = glue("ParameterPlots/abs-posterior-prediction-exp03-rect_lu-order.RDS"))
plot
ggsave("Figures/posterior-abs-exp03_rect_lu_order.png", units = "cm", width = 16, height = 10)
ggsave("Figures/posterior-abs-exp03_rect_lu_order.svg", units = "cm", width = 16, height = 10)
```

## Generating summary for predictions at the group level for relative orientation
```{r}
# Bootstrapping behavioral averages
bootstrapped_rel <- bootstrap_group_averages("BootstrappedAverages/exp3-relative-rect_lu.RDS", results, Orientation, OriResponse, compute_average_relative_response)

# Posterior predictions
posterior_rel <- posterior_group_averages_from_mu("PosteriorAverages/exp3-relative-rect_lu.RDS", results, mu, Orientation, OriResponse, compute_average_relative_response)

plot <- plot_model_predictions("rect_lu", posterior_rel, bootstrapped_rel, Rel_Stimulus, Response) +
  geom_hline(yintercept = 0, linetype = "dashed", linetype = "longdash", linewidth = 0.7)
saveRDS(plot, file = glue("ParameterPlots/rel-posterior-prediction-exp03-rect_lu.RDS"))
plot
ggsave("Figures/posterior-rel-exp03_rect_lu.png", units = "cm", width = 16, height = 10)
ggsave("Figures/posterior-rel-exp03_rect_lu.svg", units = "cm", width = 16, height = 10)
```

## Plots for posterior distributions (for weights, prior evidence relevance and uncertainty)
```{r}
prior_relevance <- parameter_posterior_df(draws, "mu_params[7]", "mu_params[8]", "Prior_Relevance")|>
  mutate("Prior Relevance" = (2 * exp(Prior_Relevance)^2))|> 
  select(.draw, Task, "Prior Relevance")
plot_parameter_distribution(prior_relevance, "exp03", parameter_name = "Prior Relevance", long_df = TRUE)

prior_relevance <- parameter_posterior_df(draws, "mu_params[7]", "mu_params[8]", "Prior_Relevance")|>
  mutate("log Prior Relevance" = log10(2 * exp(Prior_Relevance)^2))|> 
  select(.draw, Task, "log Prior Relevance")
plot_parameter_distribution(prior_relevance, "exp03", parameter_name = "log Prior Relevance", long_df = TRUE)


sigma_a <- parameter_posterior_df(draws, "mu_scale_params[5]", "mu_scale_params[6]", "Sigma_Min" )|>
  mutate("Sigma Min" = exp(Sigma_Min))
plot_parameter_distribution(sigma_a |> select(.draw, Task, "Sigma Min"), "exp03", parameter_name = "Sigma Min", long_df = TRUE)

sigma_max <- sigma_a|>
  mutate(sigma_b = parameter_posterior_df(draws, "mu_scale_params[7]", "mu_scale_params[8]", "sigma_b")[[3]],
         "Sigma Max" = exp(Sigma_Min) + exp(sigma_b))|>
  select(.draw, Task, "Sigma Max")
saveRDS(sigma_max, "ParameterDraws/sigma03.RDS")
plot_parameter_distribution(sigma_max, "exp03", parameter_name = "Sigma Max", long_df = TRUE)
ggsave("Figures/exp03-sigma_max.png", units = "cm", width = 16, height = 10)

tasks <- results["Task"] |>
  mutate(trial = glue("sigma[{1:nrow(results)}]"))
sigma_df <- as.data.frame(kilter::extract_stan_posterior_matrix(draws, "sigma"))|>
  mutate(.draw = 1:nrow(draws))|>
  pivot_longer(cols = -".draw",
               names_to = "trial",
               values_to = "Sigma") |>
  left_join(tasks, by = "trial") |>
  group_by(.draw, Task)|>
  summarise(Sigma = mean(Sigma), .groups = "drop")

plot_parameter_distribution(sigma_df, "exp03", parameter_name = "Sigma", long_df = TRUE)
```


```{r}
params03 <-
  draws |>
  select(.draw, starts_with("mu_params")) |>
  pivot_longer(starts_with("mu_params"), names_to = "Term", values_to = "mu") |>
  mutate(Index = as.integer(str_extract(Term, "\\d")),
         Task = factor((Index + 1) %% 2, labels = levels(results$Task)),
         Parameter = factor(floor((Index - 1) / 2), labels = c("Wr", "We", "Wct", "lambda"))) |>
  select(-Term, -Index)
saveRDS(params03, "ParameterDraws/params03.RDS")
```

## central task performance
```{r}
central_task <-
  results |>
  filter(Task == "dual")|>
  select(Participant, ColorTaskCorrect, ColorRT, ColorIsTarget)
  

averaging_function <- function(resample = FALSE){
  if (resample) {central_task <- slice_sample(central_task, prop = 1, replace = TRUE)}
  central_task|>
  group_by(Participant) |>
  summarise(Pcorrect = mean(ColorTaskCorrect), 
            avgRT = mean(ColorRT),
            .groups = "drop")|>
    pivot_longer(cols = c(avgRT, Pcorrect), names_to = "Parameter", values_to = "mu")
    } 

samples <- purrr::map(1:2000, ~averaging_function(resample = TRUE), .progress = TRUE) |> 
  list_rbind()

avgs <-
  samples |>
  group_by(Participant, Parameter) |>
  summarise(LowerCI = kilter::lower_ci(mu, CI = 0.97),
            UpperCI = kilter::upper_ci(mu, CI = 0.97),
            .groups = "drop") |>
  right_join(averaging_function(), by = c("Participant", "Parameter"))

saveRDS(avgs, "central_task/exp03.rds")
  
```
